---
title: "week07_JZ"
author: "Jinxiao Zhang"
date: "March 5, 2019"
output: html_document
---

```{r load library}
library(DiagrammeR)
library(rethinking)
```

1. In the Trolley data-data(Trolley)-we saw how education level (modeled as
an ordered category) is associated with responses. Is this association causal? One
plausible confound is that education is also associated with age, through a causal
process: People are older when they finish school than when they begin it.
Reconsider the Trolley data in this light. Draw a DAG that represents hypothetical
causal relationships among response, education, and age. Which statical model
or models do you need to evaluate the causal influence of education on responses?
Fit these models to the trolley data. What do you conclude about the causal relationships
among these three variables?

> Not causal association. 

```{r Q1a}
# draw the DAG
DiagrammeR("graph LR;
           Age-->Edu;
           Age-->Resp;
           Edu-->Resp")

```
> Education is not a significant predictor anymore, after considering age. Education effect changes from negative to non-significantly positive.

```{r Q1b}
data(Trolley)
d <- Trolley

levels(d$edu)
edu_levels <- c( 6 , 1 , 8 , 4 , 7 , 2 , 5 , 3 )
d$edu_new <- edu_levels[ d$edu ]

d1 <- list(
             R = d$response ,
             A = d$age, # standardize age may help make it faster!!!
             E = as.integer( d$edu_new ), # edu_new as an index
             alpha = rep(2,7) ) # delta prior
m1 <- ulam(
  alist(
    R ~ ordered_logistic( phi , kappa ),
    phi <- bE*sum( delta_j[1:E] ) + bA*A,
    kappa ~ normal( 0 , 1.5 ),
    c(bA,bE) ~ normal( 0 , 1 ),
    vector[8]: delta_j <<- append_row( 0 , delta ),
    simplex[7]: delta ~ dirichlet( alpha )
  ),
  data=d1 , chains=3 , cores=2 )

precis(m1, depth = 2, omit = "kappa") # %>% plot()



# Chapter 12
dat <- list(
             R = d$response ,
             action = d$action,
             intention = d$intention,
             contact = d$contact,
             E = as.integer( d$edu_new ), # edu_new as an index
             alpha = rep(2,7) ) # delta prior

m12.5 <- ulam(
  alist(
    R ~ ordered_logistic( phi , kappa ),
    phi <- bE*sum( delta_j[1:E] ) + bA*action + bI*intention + bC*contact,
    kappa ~ normal( 0 , 1.5 ),
    c(bA,bI,bC,bE) ~ normal( 0 , 1 ),
    vector[8]: delta_j <<- append_row( 0 , delta ),
    simplex[7]: delta ~ dirichlet( alpha )
  ),
  data=dat , chains=3 , cores=3 )


precis( m12.5 , depth=2 , omit="kappa" )

```

2. Consider one more variable in the Trolley data: Gender. Suppose that gender
might influence education as well as response directly. Draw the DAG now that
includes response, education, age, and gender.
Using only the DAG, is it possible that the inferences from Problem 1 are confounded
by gender? If so, define any additional models you need to infer the causal
influence of education on response. What do you conclude?
> Yes, a fork by gender.

```{r Q2}
DiagrammeR("graph LR;
           Age-->Edu;
           Age-->Resp;
           Edu-->Resp;
           Gen-->Edu;
           Gen-->Resp")

d2 <- list(
             R = d$response ,
             A = standardize(d$age),
             E = as.integer( d$edu_new ), # edu_new as an index
             G = d$male + 0L,
             alpha = rep(2,7)) # delta prior

m2 <- ulam(
  alist(
    R ~ ordered_logistic( phi , kappa ),
    phi <- bE*sum( delta_j[1:E] ) + bA*A + bG*G,
    kappa ~ normal( 0 , 1.5 ),
    c(bA,bE,bG) ~ normal( 0 , 1 ),
    vector[8]: delta_j <<- append_row( 0 , delta ),
    simplex[7]: delta ~ dirichlet( alpha )
  ),
  data=d2 , chains=3, cores=2 )

# show(m2)
precis(m2, depth = 2, omit = "kappa") # %>% plot()
```
> Gender has an effect. The effect of education changes from positive to slightly negative.

```{r save workspace}
save.image("week07_JZ.Rdata")
```

