---
title: "week03_JZ"
author: "Jinxiao Zhang"
date: "February 17, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(rethinking)
library(tidyverse)
```

1. Use a model to infer the total causal influence of area on weight. Would
increasing the area available to each fox make it heavier (healthier)? You
might want to standardize the variables. Regardless, use prior predictive
simulation to show that your model’s prior predictions stay within the possible
outcome range.
```{r Q1}
df.foxes <- foxes
# standardize variables
df.foxes <- df.foxes %>% 
  mutate(area_std = scale(area, scale = T),
         avgfood_std = scale(avgfood, scale = T),
         groupsize_std = scale(groupsize, scale = T))

hist(df.foxes$area) 
hist(df.foxes$weight) # range of weight: 1-8


m1 <- quap(
  alist(
    weight ~ dnorm(mu, sigma),
    mu <-  a + b*area_std,
    a ~ dnorm(4.5, 1),
    b ~ dnorm(0, 1),
    sigma ~ dexp(1)
  ), data = df.foxes
)

prior1 <- extract.prior(m1)
precis(prior1)

# plot the prior
area_seq <- seq( from=min(df.foxes$area_std) , to=max(df.foxes$area_std) ,
              length.out=50)
mu <- link( m1 , post=prior1 ,
            data=list( area_std = area_seq))

plot( NULL , xlim=range(area_seq) , ylim=c(0,10) ,
      xlab="area (std)" , ylab="weight" )
abline( h=1 , lty=2 , col = "red")
abline( h=8 , lty=2 , col = "red")
for ( i in 1:50 ) lines(area_seq , mu$mu[i,] , col=col.alpha("black",0.5) )

# for ( i in 1:N ) curve( a[i] + b1[i]*(x - xbar) + b2[i]*(x - xbar)^2,
#                         from=min(d2$weight) , to=max(d2$weight) , add=TRUE , 
#                         col=col.alpha("black",0.3) )


# evaluate the posterior
precis(m1)

post_sim1 <- sim(m1, data = df.foxes)
m1_predict <- data.frame(weight = apply(post_sim1,2,mean), area_std = df.foxes$area_std)

# plot the observed. vs. predicted
ggplot(df.foxes, aes(area_std, weight))+
  geom_point()+
  geom_point(data = m1_predict, color = "blue")
```



Answer: No, increasing the area available to each fox wouldn't make it heavier.


2. Now infer the causal impact of adding food to a territory. Would this
make foxes heavier? Which covariates do you need to adjust for to estimate
the total causal influence of food?
```{r Q2}
# Standardizeing DV makes life easier!

m2 <- quap(
  alist(
    weight ~ dnorm(mu, sigma),
    mu <-  a + bF*avgfood_std ,
    a ~ dnorm(4.5, 1),
    bF ~ dnorm(0, 1),
    sigma ~ dexp(1)
  ), data = df.foxes)

precis(m2)

```

3. Now infer the causal impact of group size. Which covariates do you need
to adjust for? Looking at the posterior distribution of the resulting model,
what do you think explains these data? That is, can you explain the estimates
for all three problems? How do they go together?
```{r Q3}
df.foxes <- df.foxes %>% 
  mutate(weight_std = scale(weight))

m3 <- quap(
  alist(
    weight ~ dnorm(mu, sigma),
    mu <-  a + bG*groupsize_std + bF*avgfood_std, 
    a ~ dnorm(4.5, 1),
    bF ~ dnorm(0, 1),
    bG ~ dnorm(0, 1),
    sigma ~ dexp(1)
  ), data = df.foxes)

precis(m3)


# From Solution
d <- foxes
d$W <- standardize(d$weight)
d$A <- standardize(d$area)
d$F <- standardize(d$avgfood)
d$G <- standardize(d$groupsize)
m3.1 <- quap(
  alist(
    W ~ dnorm( mu , sigma ),
    mu <- a + bF*F + bG*G,
    a ~ dnorm(0,0.2),
    c(bF,bG) ~ dnorm(0,0.5), # set the 2 priors together
    sigma ~ dexp(1)
  ), data=d )
precis(m3.1)
```

