---
title: "Homework_week1_JZ"
author: "Jinxiao Zhang"
date: "January 10, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

1. Suppose the globe tossing data had turned out to be 8 water in 15 tosses.
Construct the posterior distribution, using grid approximation. Use the
same flat prior as before.
```{r cars}
p_grid <- seq(from = 0, to = 1, length.out = 1000)
prior_p <- rep(1, 1000)
prob_data <- dbinom(8, 15, p_grid)
posterior <- prior_p*prob_data/sum(prior_p*prob_data)

plot(p_grid, posterior)
```
2. Start over in 1, but now use a prior that is zero below p = 0.5 and a constant
above p = 0.5. This corresponds to prior information that a majority
of the Earth's surface is water. What difference does the better prior make?
If it helps, compare posterior distributions (using both priors) to the true
value p = 0.7.

```{r}
prior_p2 <- c(rep(0, 500),rep(1,500))
prob_data2 <- dbinom(8, 15, p_grid)
posterior2 <- prior_p2*prob_data2/sum(prior_p2*prob_data2)

plot(p_grid, posterior2)
points(p_grid, posterior, col = "grey")

# for (i in 1:1000){
#  if (posterior2[i] == max(posterior2)) {
#    print(i)
#    break
#  }
# }

```

3. This problem is more open-ended than the others. Feel free to collaborate
on the solution. Suppose you want to estimate the Earth's proportion of
water very precisely. Specifically, you want the 99% percentile interval of the
posterior distribution of p to be only 0.05 wide. This means the distance between
the upper and lower bound of the interval should be 0.05. How many
times will you have to toss the globe to do this? I won't require a precise
answer. I'm honestly more interested in your approach.

```{r}
for (n in seq(1000, 10000, by = 10)){
  prob_data3 <- dbinom(n*0.7, n, p_grid)
  posterior3 <- prior_p*prob_data3/sum(prior_p*prob_data3)
  
  #plot(p_grid, posterior3)
  
  temp1 = 0
  for (i in 1:(n/2)){
    temp1 = temp1 + posterior3[i]
    if (temp1 >= .005){
      print(c(n,i))
      break
    }
  }
  
  temp2 = 0
  for (j in 1:(n/2)){
    temp2 = temp2 + posterior3[1001-j]
    if (temp2 >= .005){
      print(c(n,1001-j))
      break
    }
  }
  
  if((1001-j-i) <= 50){
    message(paste("found it: ",n))
    answer = n
    break
  }
  
}

```

